### 2011-11-11 papers/lmcproc/lmcproc.tex
Davide's comments added
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@1481 40876441-3496-0410-a613-a2fca35b2ae2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt,twoside]{iopart}
\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
\usepackage{iopams} 
\usepackage{graphicx, subfigure}
\usepackage{balance}
\usepackage{bm}
\usepackage{float}
\newcommand{\ex}[1]{\times10^{#1}}

\begin{document}

\title[]{Confined Cubic Blue Phases under Shear}
\author{O. Henrich$^{1}$, K. Stratford$^2$, D. Marenduzzo$^3$, P.V. Coveney$^1$ and M.E. Cates $^3$}

\address{$^1$ Centre for Computational Science, University College London, UK\\$^2$ Edinburgh Parallel Computing Centre, University of Edinburgh, UK\\$^3$ School of Physics and Astronomy, University of Edinburgh, UK}
\ead{o.henrich@ucl.ac.uk}

\begin{abstract}
We study the behaviour of confined cubic blue phases under a shear flow, via lattice Boltzmann simulations. We focus on the two experimentally observed
phases, blue phase I and blue phase II. The disclination network of blue phase II continuously breaks and reforms under shear, 
leading to an oscillatory stress response in time. The oscillations are only regular for very thin samples. For thicker samples, the shear leads to a ``stick-slip'' motion
of part of the network along the vorticity direction. Blue phase I responds vastly differently: its defect network undergoes seemingly
chaotic rearrangements under shear, irrespective of system size. 
\end{abstract}

%Uncomment for PACS numbers title message
%\pacs{00.00, 20.00, 42.10}
% Keywords required only for MST, PB, PMB, PM, JOA, JOB? 
%\vspace{2pc}
%\noindent{\it Keywords}: Article preparation, IOP journals
% Uncomment for Submitted to journal title message
%\submitto{\JPA}
% Comment out if separate title page not required
%\maketitle

\section{Introduction}
% Put \label in argument of \section for cross-referencing
%\section{\label{}}
Cholesterics are liquid crystals in which the local director field spontaneously twist in thermodynamic equilibrium~\cite{deGennes}.
The preferred configuration close to the isotropic boundary features twist around two perpendicular axes, as opposed to just one axis in the regular cholesteric state, and the corresponding deformation is denoted a ``double-twist cylinder''.
As it is topologically impossible to continuously patch 3D space with double-twist cylinders, defects arise. These may organise into a variety of regular periodic lattices, giving rise to the so-called cubic blue phases (BPs) \cite{Grebel:1984,Wright:1989}. There are two experimentally observed cubic blue phases, BPI and BPII (a third, BPIII, is likely amorphous~\cite{Henrich:2011a}).\\
BPs have been long considered as purely of academic interest due to their very narrow range of stability, but this has drastically changed since the creation of polymer- and temperature-stabilised BPs~\cite{Kikuchi:2002,Coles:2005}, which has opened up the possibility of novel applications.
During the last few years considerable progress has been achieved regarding the behaviour of BPs in confined geometries~\cite{Fukuda:2010a, Fukuda:2010b, Ravnik:2011b}, under external fields \cite{Alexander:2008,Fukuda:2009,Henrich:2010a,Castles:2010,Tiribocchi:2011}, and in the presence of colloidal particles \cite{Ravnik:2011a}. 
The kinetics of BP domain growth have been recently addressed in~\cite{Henrich:2010b}. However, our understanding of their dynamical behaviour under flow remains
very limited. The aim of this work is to contribute to fill this gap by studying, for the first time, the response of confined BP samples to a shear flow. \\
Flow in cholesterics is both strongly non-Newtonian and highly anisotropic. If a cholesteric helix is subjected to, for instance, a Poiseuille flow along the helical axis, for small pressure differences flow occurs mainly through ``permeation'', first investigated by Helfrich \cite{Helfrich:1969}. In the permeation mode the liquid crystal flows by leaving the director field virtually unchanged, and this leads to high dissipation and large viscosities. Marenduzzo et al. \cite{Marenduzzo:2006a,Marenduzzo:2006b} simulated shear and Poiseuille flow in cholesteric liquid crystals in the permeation mode, and showed the importance of the boundary conditions in determining the apparent viscosity of the fluid. They also found that a strong secondary flow appears.
Rey \cite{Rey:1996a, Rey:1996b} studied shear in cholesterics oriented with the helix along the vorticity axis and found that at low Ericksen number travelling twist waves appear which lead to the rotation of the cholesteric helix. At higher forcing, the helix uncoils and leaves a flow-induced nematic phase. 
Rey also studied cholesterics subjected to steady and low frequency small amplitude oscillatory shear for different helix orientations \cite{Rey:2000, Rey:2002}, where he found that splay–bend–twist deformations were excited when the helix is aligned along the flow direction. Splay–bend deformation occured when the helix is aligned along the velocity gradient, and only twist deformations when the helix is aligned along the vorticity axis. \\
Dupuis et al. \cite{Dupuis:2005} performed the first numerical investigation of BP rheology in Poiseuille flow, starting from equilibrium structures of BPI and BPII and a periodic array of doubly twisted cylinders.
Under small forcing, the network opposed the flow giving rise to a significant increase in apparent viscosity.
Upon increasing the forcing they found clear evidence of shear thinning.
In the crossover region they predicted a novel oscillatory regime where the network continuously breaks and reforms as portions of the disclinations in the center of the channel move to neighboring cells and relink with the parts of the network left behind by the flow. With respect to the cholesteric case, the viscosity still decreases with forcing (the system shear thins) but much less than in cholesterics in the permeation mode, which is in agreement with experiments experiments \cite{Zapotocky:1999, Ramos:2002}.\\
This paper is organised as follows: 
In Sec. \ref{sec_model} we introduce the Beris-Edwards model for hydrodynamics of liquid crystals, which is a generalisation of the Eriksen and Leslie's theory of nematodynamics~\cite{deGennes}.
In Sec. \ref{sec_results} we describe the simulation methodology and boundary conditions we applied.
Results of a confined BPII in rectilinear shear flow are presented in Sec. \ref{sec_bpii} with particular focus on the dependence of the thickness of the sample and thermodynamic state.
In Sec. \ref{sec_bpi} we show corresponding results for BPI, which indicates fundamental differences in the flow behaviour of BPI and BPII.

\section{Model}\label{sec_model}

Our approach is based on a well-established model for the hydrodynamics of cholesteric liquid crystals \cite{Beris:1994,Olmsted:1999} that describes their state in terms of a traceless, symmetric tensor order parameter ${\mathbf Q}({\mathbf r})$. 
In the uniaxial approximation the order parameter is given by $Q_{\alpha \beta}=\langle n_\alpha n_\beta - 1/3\; \delta_{\alpha\beta}\rangle$ with ${\mathbf n}$ the director field.
Brackets indicate mesoscopic averaging over a large number of individual molecules.
The largest eigenvalue of ${\mathbf Q}$, $0\le q_s\le\frac{2}{3}$ characterizes the local degree of orientational order.
The thermodynamical properties of the liquid crystal are determined by a free energy
${\cal F}$, whose density $f$ consists of a bulk contribution $f_b$ and a gradient part $f_g$, as follows,
\begin{eqnarray}
f_b&=&\frac{A_0}{2}\left(1-\frac{\gamma}{3}\right) Q_{\alpha \beta}^2-\frac{A_0 \gamma}{3}Q_{\alpha \beta} Q_{\beta \gamma} Q_{\gamma \alpha}+\frac{A_0 \gamma}{4}(Q_{\alpha \beta}^2)^2\nonumber\\
f_g&=&\frac{K}{2}(\varepsilon_{\alpha\gamma\delta} \partial_\gamma Q_{\delta\beta}+2 q_0 Q_{\alpha \beta})^2+\frac{K}{2}(\partial_\beta Q_{\alpha \beta})^2\label{FE}
\end{eqnarray}
The first term contains the bulk-free energy constant $A_0$ and the inverse temperature $\gamma$ which controls the magnitude of order.
The second part quantifies the cost of elastic distortions, which are proportional to the elastic constant $K$ --
we work for simplicity in the one-elastic constant approximation. The wavevector $q_0=2\pi/p_0$, where $p_0$ is the cholesteric pitch.
The actual periodicity of the BP structure, $p$, does not need to be equal to $p_0$.
The ``redshift'' $r=p/p_0$ is adjusted during the simulation by following a simple procedure similar to \cite{Alexander:2006}.\\
A particular thermodynamic state is specified by two dimensionless quantities: the effective temperature $\tau$ and chirality $\kappa$,
which are given by
\begin{eqnarray}
\tau&=&\frac{27(1-\gamma/3)}{\gamma}\nonumber\\
\kappa&=&\sqrt{\frac{108 K q_0^2}{A_0 \gamma}}\nonumber.
\end{eqnarray}

The dynamical evolution of the order parameter is given by the equation 
\begin{equation}
\left(\partial_t+ u_\alpha \partial_\alpha \right){\mathbf Q} - {\mathbf S}({\mathbf W},{\mathbf Q}) = \Gamma {\mathbf H}.
\label{eqn2}
\end{equation}
The first term on the left hand side of Eq.\ref{eqn2} is a material derivative, which describes the rate of change of a quantity moving along with the flow.
The second term accounts for the rate of change due to local velocity gradients $W_{\alpha \beta}=\partial_\beta u_\alpha$,
\begin{eqnarray}
{\mathbf S}({\mathbf W}, {\mathbf Q}) &=& (\xi {\mathbf A} + {\boldsymbol \Omega})({\mathbf Q}+\frac{\mathbf I}{3})\nonumber\\
& &\hspace*{-1.5cm}+ ({\mathbf Q}+\frac{\mathbf I}{3})(\xi {\mathbf A}  - {\boldsymbol \Omega})-2 \xi ({\mathbf Q}+\frac{\mathbf I}{3})Tr({\mathbf Q W}),
\label{eqn3}
\end{eqnarray}
where $Tr$ denotes the tensorial trace and ${\mathbf A}=({\mathbf W}+{\mathbf W}^T)/2$ and ${\boldsymbol \Omega}=({\mathbf W}-{\mathbf W}^T)/2$ are the symmetric and antisymmetric part of the velocity gradients, respectively. $\xi$ is a constant depending on the molecular details of the liquid crystals.
Flow alignment occurs if $\xi \cos{2\theta}=\frac{3q_s}{2+q_s}$ has a real solution, where $\theta$ is the Leslie-angle: we select this case by setting $\xi=0.7$ in our runs.
${\mathbf H}$ is the molecular field, which is a functional derivative of $\cal F$ that preserves the tracelessness of $\mathbf Q$.
\begin{equation}
{\bf H}=-\frac{\delta {\cal F}}{\delta {\bf Q}}+\frac{\bf I}{3}\, Tr\left(\frac{\delta {\cal F}}{\delta {\bf Q}}\right).
\label{eqn4}
\end{equation}
The rotational diffusion constant $\Gamma$ is proportional to the inverse of the rotational viscosity $\gamma_1=2 q_s^2/\Gamma$.\\
The time evolution of density and velocity are respectively governed by the continuity and the following 
Navier-Stokes equation Eq.\ref{eqn6}, 
\begin{eqnarray}
\partial_t u_\alpha +\rho \,u_\beta \partial_\beta u_\alpha&=&\partial_\beta \Pi_{\alpha \beta}+\eta\, \partial_\beta \{ \partial_\alpha u_\beta + \partial_\beta u_\alpha +(1+3\frac{P_0}{\partial \rho} )\partial_\mu u_\mu \delta_{\alpha \beta})\}. 
\label{eqn6}
\end{eqnarray}

At low flow rates the fluid can be considered as incompressible, so that the last term on the righthand side of Eq.\ref{eqn6} can be safely neglected.
$\eta$ is an isotropic background viscosity which is set to $\eta=0.8333$ in our case.
The pressure tensor reads explicitly
\begin{eqnarray}
\Pi_{\alpha \beta}&=&P_0-\xi H_{\alpha \gamma}\left(Q_{\gamma \beta} +\frac{1}{3} \delta_{\gamma \beta}\right)-\xi \left(Q_{\alpha \gamma} +\frac{1}{3} \delta_{\alpha \gamma}\right) H_{\gamma \beta}\nonumber\\
&+& 2 \xi  \left(Q_{\alpha \beta} +\frac{1}{3} \delta_{\alpha \beta}\right) Q_{\gamma \nu} H_{\gamma \nu}-\partial_\alpha Q_{\gamma \nu} \frac{\delta{\cal F}}{\delta \partial_{\beta} Q_{\gamma \nu}}\nonumber\\
&+&Q_{\alpha \gamma}H_{\gamma \beta}-H_{\alpha \gamma} Q_{\gamma \beta}.
\label{eqn7}
\end{eqnarray}
In the isotropic state when ${\bf Q}\equiv 0$ Eq.\ref{eqn7} is reduced to the scalar pressure, which is constant to a very good approximation.\\ 
The system of coupled partial differential equations is solved by means of a hybrid scheme \cite{Marenduzzo:2007} that uses a lattice Boltzmann algorithm with predictor-corrector scheme for the continuity equation and Eq.\ref{eqn6} and a finite difference scheme for the equation of motion of the tensor order parameter Eq.\ref{eqn2}.
More details on the algorithm can be found in \cite{Denniston:2001, Denniston:2004}.

\section{Methodology and Results}\label{sec_results}

In the following we present results from simulations of confined cubic blue phases in steady shear flow.
Our system consists of $n_y\times n_z = 4\times4$ unit cells in y- and z-direction (flow and vorticity direction respectively) and a varying number $n_x$ of unit cells along a gap in x-direction (which is the flow gradient direction in our geometry).
As is common in BP simulation work~\cite{Henrich:2010b,Henrich:2011a}, we initialised our simulations with analytical solutions that minimize the free energy functional Eq.\ref{FE} in the high-chirality limit, and equilibrated these (for 3000 LB timesteps) with periodic boundary conditions prior to starting the shear. During the equilibration run the optimal redshift $r$ was recalculated
at every timestep.

Shear flow was then started by applying shear boundary conditions to the LB distribution functions as in~\cite{Denniston:2004}. As regards ${\mathbf Q}$, the equilibrated order parameter was advected on the top and bottom boundary, according to the imposed wall velocities. Periodic boundary conditions were retained along the flow direction and the direction of vorticity.
This set-up is similar to the one Dupuis et al. used to simulate Poiseuille flow \cite{Dupuis:2005}. The pinning of the defect network at the boundaries may be realised in practice by impurities or surface defects.
During shear flow the unit cell size was kept fixed and a constant redshift equal to the value at the end of the equilibration protocol was assumed.\\ 
The total number of unit cells in our simulations was $n_x\times n_y \times n_z= 1, \dots, 4 \times 4 \times 4$ for both blue phases, using a resolution of 32 lattice Boltzmann units (LBU) per pitch length (16 per unit cell) for BPII and 64 LBU per pitch length (32 per unit cell) for BPI.  
This provided enough spatial resolution to track even complex rearrangements of the disclination networks.
A timestep and lattice spacing in LBU corresponds to $\sim 1 {\rm ns}$ and $\sim 10{\rm nm}$ in SI-units.
The LBU of stress equals to about $10^8 Pa$. 
More details about the conversion from LBU to SI-units can be found in \cite{Henrich:2010b,Henrich:2011a}.

\subsection{Blue Phase II}\label{sec_bpii}

\begin{figure}[t]
\centering
\includegraphics[width=0.24\textwidth]{bp2_t-0o5_k2_25k_y.png}
\includegraphics[width=0.24\textwidth]{bp2_t-0o5_k2_50k_y.png}
\includegraphics[width=0.24\textwidth]{bp2_t-0o5_k2_75k_y.png}
\includegraphics[width=0.24\textwidth]{bp2_t-0o5_k2_100k_y.png}
\caption{Shear flow of BPII at different gap sizes: The pictures show a section of the entire disclination network of $n_x=1 \mbox{ (yellow)}, 2 \mbox{ (green)}, 3 \mbox{ (red) and } 4 \mbox{ (blue)}$ unit cells along the gap and $n_y \times n_z =2 \times 2$ unit cells in flow and vorticity direction. The thermodynamic parameters temperature and chirality are $\tau=-0.5$ and chirality $\kappa=2$, respectively, where BPII is the equlibrium phase. The velocity gradient is oriented along the vertical direction with the top (bottom) wall moving into (out of) the viewing plane. The strain rate is $\dot{\gamma}=7.8125\ex{-5}$ for all gap sizes.}
\label{fig1}
\end{figure}

We first report the results for BPII.
Fig.\ref{fig1} shows the disclination network in steady shear flow for different number of unit cells between the walls.
(We prefer the perspective along the flow direction because the network topology is clearer then.)
Temperature and chirality are $\tau=-0.5$ and $\kappa=2$ respectively, and for these parameters BPII is the equilibrium phase.
The velocity of the walls has been adapted so that the same shear rate $\dot{\gamma}=7.8125\ex{-5}$ is achieved for all gap sizes.
Shear rates four times larger and smaller were also applied and produced very similar results, apart from the flow velocities which scale up accordingly.
For all gap sizes we observe an affine transformation in gradient-flow plane that leads to a break-up of the network at strains of about half a unit cell size.
Shorlty after the break-up the network reforms further downstream, resulting in a periodic, regular structure that is very close to the equilibrium
BPII disclination network. Similar oscillations have been previously reported for
BPs in Poiseuille flow, with a gap of one unit cell size \cite{Dupuis:2005}.\\
It is interesting to compare the results of different gap sizes.
For $n_x=1$ we observe that the BPII-network is slightly deformed with respect to the quiescent state, but reforms always at the same position.
At gap sizes of more than one unit cell the BPII network distorts more under the same shear rate, 
suggesting that the structure within the sample is highly affected by the anchoring at the boundaries.
Interestingly, when the sample is thicker than one unit cell a movement of the network in negative vorticity direction arises, 
which is not observed for $n_x=1$.
This network motion does not proceeds at a constant speed, rather it proceeds in a ``stick-slip'' fashion: elastic stresses first slowly build up, and are then 
quickly released as the network ``slips'' along the vorticity direction.
Changing from a left- to a right-handed helix inverts the direction of motion of the network.
Hence the ``stick-slip'' motion is related to the chiral nature of the helix. This phenomenon may also bear some similarities to
the travelling wave motion in cholesterics sheared along the vorticity direction 
below a critical Ericksen number \cite{Rey:1996a, Rey:1996b}.
It is interesting to note that the secondary flow in the BP
does {\em not} match the direction of the ``stick-slip'' motion of the network, which is then not connected to mass flow
(in this respect this phenomenon is close to a permeation flow).

\begin{figure*}[t]
\centering
\includegraphics[width=0.3\textwidth]{disc_bp2_tk_scan_25k.png}
\includegraphics[width=0.3\textwidth]{disc_bp2_tk_scan_75k.png}
\includegraphics[width=0.3\textwidth]{disc_bp2_tk_scan_125k.png}
\caption{BPII in shear flow for different thermodynamic parameters: The images show the network as it appears along the flow direction at the reference state (blue, $\tau=-0.5, \kappa=2$), a high temperature state (red, $\tau=0.5, \kappa=2$) and a low chirality state (green, $\tau=-0.5, \kappa=1$), where BPII is metastable. The strain rate is $\dot{\gamma}=7.8125\ex{-5}$ in all three cases. The pictures show a central section consisting of 8 unit cells from a larger run with $n_x=4$ gap size.}
\label{fig2}
\end{figure*}

The speed of the network in vorticity direction depends on the thermodynamic state of the blue phase.
A comparison between the state from Fig. \ref{fig1} with a ``high temperature'' state at $\tau=0.5, \kappa=2$ and a ``low chirality state'' at $\tau=-0.5, \kappa=1$
is shown in Fig.\ref{fig2}. Note that in both these states BPII is now a metastable phase. 
The images in Fig.~\ref{fig2} give a sequence of initially overlapping disclination networks that separate during the course of the simulation.
Of all three states the high temperature state is the fastest one, followed by the reference state and the low chirality state.

\begin{figure*}[t]
\centering
\includegraphics[width=0.75\textwidth]{stress_bp2_fbc.pdf}
\caption{BPII: Ensemble-averaged shear stress $\langle \Pi_{xy}\rangle$ vs. time in the reference state ($\tau=-0.5, \kappa=2$) for gaps ranging from one to four unit cells across.}
\label{fig3}
\end{figure*}

The total network displacement along the vorticity direction is more than a unit cell within $10^5$ timesteps, for the ``high-temperature'' state.
Typical average and maximum values of the scalar order parameter are $\bar{q}_s=0.338, q_{s,max}=0.383$ (low chirality state), 
$\bar{q}_s=0.206, q_{s,max}=0.320$ (high temperature state) and $\bar{q}_s=0.304, q_{s,max}=0.408$ (reference state, in Fig.~\ref{fig1}).
The average order parameter of the ``low-chirality'' state is by about $10\%$ higher compared to the reference state, 
whereas the relative differerence between the mean and the maximum value is about a factor $2.5$ smaller.
We propose that the speed of the networks is different because permeation, which should be required for the ``stick-slip'' motion, may occur more easily in states with less order (``high-temperature'' state) and gentler spatial variation (``low-chirality'' state). Furthermore, the uneven, ``stick-slip'' nature of the network motion is possibly due to the anchoring at the boundaries which prevents network displacements close to there.

Fig. \ref{fig3} shows the spatially averaged shear stress $\langle \Pi_{xy}\rangle$ versus time, in the reference state and for different gap sizes.
For gaps accommodating only one unit cell a regular periodic pattern emerges shortly after a transient start-up phase.
Every cycle is related to a break-up and reformation of the network.
The time-average in the steady state is about $\Pi_{xy}\simeq5.62\ex{-5}$ in LBU and yields an apparent viscosity $\eta_{app}=(1 +\Pi_{xy}/\eta\dot{\gamma})\simeq 1.86$. 
When the gap contains $n_x\ge2$ unit cells, the average stress shows less regular oscillations, and fluctuates around a mean value which is about $10-15\%$ smaller, roughly $\eta_{app}=1.65$ for $n_x=4$. The fact that the average stress stabilises for $n_x\ge2$ indicates that bulk flow behaviour should occur for those thicknesses, but not in the single unit cell case.
Different shear rates lead to similar generic features.

\subsection{Blue Phase I}\label{sec_bpi}

We now compare and contrast the results just obtained with those pertaining to the rheology of blue phase I. The topology of the disclination network in BPI in equilibrium is very different from that of BPII: the main difference is the lackin BPI of junction points where several (four in BPII) disclination lines of topological charge -1/2 meet. \\
In the following we give representative results for BPI at $\tau=-0.5$ and $\kappa=1$, where it is the equilibrium phase.
Fig.\ref{fig4} shows the disclination network at timestep $t=4\ex{5}$ for different gap sizes, with the viewing direction again chosen along the flow direction.
For clarity we only show $n_x \times n_y \times n_z=1,\dots4\times1\times2$ unit cells.

\begin{figure*}[h]
\centering
\begin{minipage}[b]{0.3\textwidth}
\subfigure[]{\includegraphics[width=\textwidth]{disc_bp1_t-0o5_k1o0_400k_1uc.png}}
\subfigure[]{\includegraphics[width=\textwidth]{disc_bp1_t-0o5_k1o0_400k_2uc.png}}
\end{minipage}
\subfigure[]{\includegraphics[width=0.3\textwidth]{disc_bp1_t-0o5_k1o0_400k_3uc.png}} 
\subfigure[]{\includegraphics[width=0.3\textwidth]{disc_bp1_t-0o5_k1o0_400k_4uc.png}}
\caption{Confined BPI in shear flow at $\tau=-0.5, \kappa=1$ and and gap sizes of one (a, yellow), two (b, green), three (c, red) and four (d, blue) unit cells. The pictures show the defect network along the flow direction at timestep $t=4\ex{5}$ at the same flow rate as above.}
\label{fig4}
\end{figure*}

\begin{figure*}[h]
\centering
\includegraphics[width=0.75\textwidth]{stress_bp1_fbc.pdf}
\caption{BPI: Ensemble-averaged shear stress $\langle \Pi_{xy}\rangle$ vs. time in the reference state ($\tau=-0.5, \kappa=1$) for gaps ranging from one to four unit cells across.}
\label{fig5}
\end{figure*}

Contrary to BPII, the disclination network of BPI changes greatly under a shear flow, and the typical snapshots in Fig.~\ref{fig4} bear no relation to the equilibrium BPI configuration. Furthermore, there is no sign of the ``stick-slip'' motion of the defects observed in BPII. While no regular cubic structure emerges under flow, it is interesting to note that the network develops a periodic modulation along the vorticity direction (horizontal in Fig.~\ref{fig4}). This periodicity arises for all gap sizes. Whether this feature is a consequence of frustration with the anchoring at the boundaries, or is a prelude to a slow ordering of the network under flow, we cannot yet tell.

The ensemble averages of the shear stress $\Pi_{xy}$ for different gap sizes is shown in Fig.\ref{fig5} and gives more evidence of the chaotic nature of the flow of BPI.
The magnitude of the average shear stress as well as the drop when increasing the gap size seem are both comparable to those observed with BPI.

\section{Discussions and Conclusions}

In conclusion, we presented the first supra-unit cell simulations of confined cubic blue phases in simple shear flow.
The rheological behaviour of BPI and BPII was found to be strikingly different.

Our simulations show that the disclination network of BPII continuously breaks and reforms while the network undergoes an affine deformation under shear. This process leads to an oscillatory shear stress in time, although the oscillations are only regular for gap sizes of a single unit cell across the flow gradient directions, where anchoring affects the physics of the system greatly, in agreement with recent simulations of quiescent BP networks \cite{Fukuda:2010a, Fukuda:2010b, Ravnik:2011b}.
For larger gap sizes of two and more unit cells a ``stick-slip'' motion of the network arises, along the vorticity direction. The associated flow is not directly linked to mass flow, and is therefore somewhat akin to permeation flow in cholesterics. The average speed of this network translation depends on the thermodynamic parameters, and appears to be higher for higher temperature and lower chirality.

Contrary to BPII, BPI shows no regular break-up and relinking of the unit cells and its disclination network undergoes chaotic rearrangements under shear.
This behaviour was found for all gap sizes, and may be linked to the largely different topology in the equilibrium disclination patterns in BPI and BPII.
We will adress the rheology of BPI more in detail in an upcoming publication \cite{Henrich:2012}.

% If you have acknowledgments, this puts in the proper section head.
\ack
We acknowledge support by EPSRC Grants No. EP/E045316 and No. EP/E030173 and computing time on HECToR, U.K.
M.E.C. holds a Royal Society Research Professorship.

% Create the reference section using BibTeX:

\section*{References}

\bibliography{lmcproc}
\bibliographystyle{unsrt}
\end{document}