### 2014-02-20 trunk/papers/2012-sandwich/response.txt
Reviewers comments and repsonse
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2310 40876441-3496-0410-a613-a2fca35b2ae2

Date: Mon, 17 Feb 2014 04:37:57 -0500
From: c.huang@nature.com
To: dmarendu@ph.ed.ac.uk
Subject: Decision on manuscript NCOMMS-14-00180-T

** Please ensure you delete the link to your author home page in this e-mail
if you wish to forward it to your coauthors **

Dear Dr Marenduzzo,

Your manuscript entitled "Self-assembly of colloid-cholesteric composites: A
route to switchable optical materials" has now been seen by 2 referees,
whose comments are appended below. You
will see that, while they find your work of interest, they have raised
points that need to be addressed before we can make a decision on
publication.

We therefore invite you to revise and resubmit your manuscript, taking into
account the points raised. At the same time, we ask that you ensure your
manuscript complies with our format
requirements, which are summarized in the following checklist:
http://www.nature.com/ncomms/authors/ncomms_manuscript_checklist.pdf

Please use the following link to submit your revised manuscript, completed
checklist and point-by-point response to the referees' comments (which
should be in a separate document to any
cover letter):
http://mts-ncomms.nature.com/cgi-bin/main.plex?el=A6S2Lka4A1DfBL5I4A9v6k5dsV
e8yORLGQzsElBAZ

** This url links to your confidential home page and associated information
about manuscripts you may have submitted or be reviewing for us. If you wish
to forward this email to co-authors,
please delete the link to your homepage first **

We hope to receive your revised paper within two months; please let us know
if you aren't able to submit it within this time so that we can discuss the
best way forward. If we don't hear
from you, and the revision process takes significantly longer, we will close
your file. In this event, we will still be happy to reconsider your paper at
a later date, as long as nothing
similar has been accepted for publication at Nature Communications or
published elsewhere in the meantime.

We look forward to hearing from you soon.

Best regards,

Congcong Huang, PhD

Associate Editor
Nature Communications

Unit 10-11, 42F, The Center
989 Changle Road, Xuhui District
Shanghai, China, 200031
Tel: +86-21-2422-5011
Fax: +86-21-2422-5015
www.nature.com/naturecommunications

Reviewers' comments:

Reviewer #1 (Remarks to the Author):

The authors describe the results of a study of colloidal particles dispersed
in cholesteric blue phases (BP) focusing on the effects of particle
concentration and distribution. Additionally
effects of confinement and external fields are investigated. The use of the
lattice Boltzmann approach allows to follow dynamics after the thermal
quench and electric field application. The
study is an important step in modeling of BPs and should deserve a
publication in Nature Communications. Nevertheless there are some weak
points that need improvement before the publication.
Below is the list of remarks.

General remarks:

The term colloid is by the definition a dispersion of particles in a fluid.
Using the term also for the particles my lead to a confusion. In the
manuscript often also a more appropriate word
colloidal particle is used.

It would be important to add comments that would help experimentalist in
designing experiments with colloidal blue phases and perhaps stimulate
development of applications on the base of
these materials.

Particular remarks:

Page 2

In the discussion of the physical understanding of trapping of colloidal
particles in BP disclination networks it would be useful to refer to the
initial study by Ravnik et al. in Faraday
Discussions 2009 devoted to this topic.

250 K is more than the total mesophase stability range!

Page 3

The statement "This is defined such that W > 0 gives preferential
perpendicular orientation of the director field at the particle surface." is
not consistent with the interaction form
described in the Supplement.

For potential readers it would be useful to specify the particle radius in
nanometers in the main text. How it compares to the nematic correlation
length? How the radius influences the
structures? How tiny the particles can still be that the method based on the
Landau-de Gennes free energy is justified.

Is WR/K = 2.8 justified for nanoparticles used in the study?

Page 4:

Can the G elastic modulus be estimated for modeled structures?

Page 5:

Can annealing of the structures obtained with the thermal quench lead to
ordered colloidal structures?

Is also in case of the disordered structures observed a change in the
stability range of the colloidal BPs?

In the discussion of colloidal particles in BP thin layers it would be good
to comment the study of
Ravnik et al. in Soft Matter 2011 devoted the ordered structures of such a
system.

Page 6:

It would be good to comment the difference between structures in the bulk
and thin layer for particles with strong anchoring. Does this depend on the
visualization or is more behind?

Are the electric field induced changes irreversible only in case of
colloidal dispersions?

Page 7

Reference 2 has a mistake.

Supplement

Segment 4.2:

The statement "Wall anchoring is either normal or planar with W=K = 1 in
simulation units" is unclear.

References of the Supplement:
There are mistakes in references 1 and 2.

Reviewer #2 (Remarks to the Author):

This paper, "Self-assembly of colloid-cholesteric composities: A route to
switchable optical materials" reports simulations of the formation of
particle networks in a cholesteric host and in
particular in the blue phase of the cholesteric host. It reports the
influence of various physical parameters and in particular the surface
anchoring and concentration of the colloidal
particles. The paper also reports how external fields can be employed to
change the morphology of these colloidal composites and how this may be
employed in practical optical devices.

The paper builds on a body of previous research, both theoretical and
experimental that has explored the effects of dispersing particles in a
liquid crystal host. The novelty of the paper is
to model the effect of particles in the blue phases of a cholesteric host
ahd to postulate how these colloids may be employed in specific
appilcations.

I am not a theorteical physiscist and therefore will not comment on the
methodology but rather will take the perspective of an experimentalist and
how useful this paper is in understanding
these materials and more importantly how useful the paper is in suggesting
and designing experimental procedures and developing the actual materials.

The paper provides insight into the effectiveness of including particles
into the blue phase as a means of controlling and adjusting the structure
and stability of the materials. The paper
is clearly written and the figures, paricularly the included 3D videos,
provide clear perspective on the modeled phases of these materials. The
results are of interest and useful to
experimentalists working in the field. The paper would be even more useful
if it included references to actual materials, both the liquid crystal host
and the added particles and perhaps
even suggested a model system that would be a good starting point to study
experimentally. In a similar vein, it would also be useful to address the
practical issue of reproducibility, which
has been a major problem with these materials. The theoretical results
clearly indicate how important the formation conditions are to the final
result. What is the authors sense of how easy
it will be to reproducibly fabricate a device
with a specific morphology and optical properties. This seems a likely
practical issue particularly with the rich variety of structures reported in
the simulations.

** See NPG's author and referees' website at www.nature.com/authors for
information about policies, services and author benefits

This email has been sent through the NPG Manuscript Tracking System
NY-610A-NPG&MTS

Confidentiality Statement:

This e-mail is confidential and subject to copyright. Any unauthorised use
or disclosure of its contents is prohibited. If you have received this email
in error please notify our Manuscript
Tracking System Helpdesk team at http://npgapps.custhelp.com/app/ask.

Details of the confidentiality and pre-publicity policy may be found here
http://www.nature.com/authors/policies/confidentiality.html

Privacy Policy | Update Profile