### 2014-02-20 papers/2012-sandwich/reply.tex
Added draft reply
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@2311 40876441-3496-0410-a613-a2fca35b2ae2

%\documentclass[aps,prl,groupedaddress,showpacs]{revtex4}
\documentclass[12pt]{article}

\usepackage{graphicx,subfigure}
\usepackage{color}

\pagestyle{empty}

\begin{document}

Dear Editor,\\

We would like to thank you and the referees for valuable comments to our 
manuscript, `Self-assembly of colloid-cholesteric composites: A 
route to switchable optical materials'. \\

Please find below a summary of the changes we applied and a detailed
response to the comments of the referees:\\

{\bf Referee\#1}\\

{\it General remarks}\\
- The term colloid has been replaced with colloidal particle throughout the main body of the manuscript and the supplement.\\
- Added comment about experimental realisations of BPs.\\

{\it Particular remarks}\\
- Ravnik et al Faraday Disc. 2009 has been cited.\\
- 250K and stability range of the mesophase\\
- inconsistency of the interaction form between supplement and main article\\
- specify particle radius and validity range of Landau-de Gennes FE functional\\
- WR/K=2.8 is strong anchoring regime\\
- elastic modulus G\\
- Annealing and thermal quench does not lead to ordered structures as can be seen from Fig.\\
- We did not determine the stability range of the disordered colloid-BP structures.\\
- A comment and reference to Ravnik et al. Soft Matter 2011 has been added.\\
- Comment on difference between structures in the bulk and thin layers with strong anchoring\\
- The changes induced by the electric field are only irreversible in the presence of colloidal particles with sufficiently strong anchoring.\\
- Reference 2 has been corrected\\

{\it Supplement}\\
- The comment about planar and normal wall anchoring has been changed and clarified.\\
- References 1 and 2 have been corrected.\\

{\bf Referee\#2}\\

{\it General remarks}\\
- References to actual materials, liquid crystal host materials and particles have been added.\\
- Comment about reproducibility of these materials has been added.\\

\end{document}