### 2014-03-04 trunk/papers/2012-sandwich/response.txt
updated response (Davide)
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2324 40876441-3496-0410-a613-a2fca35b2ae2

Dear Dr Huang,

many thanks for sending us the referee reports on our manuscript, 
"Self-assembly of colloid-cholesteric composites: A route to switchable 
optical materials". We are glad that both referees found our work interesting.
We are also grateful for the suggestions for improvement, all of which
we have now implemented in the revised version. We hope the revised
manuscript can now be accepted for publication in Nature Communication.

With best wishes,

K. Stratford, O. Henrich, J. S. Lintuvuori, M. E. Cates and D. Marenduzzo

------------------------------------------------------------------

Point by point response to referees' comments

Reply to Reviewer #1:

COMMENT:
"The authors describe the results of a study of colloidal particles dispersed
in cholesteric blue phases (BP) focusing on the effects of particle
concentration and distribution. Additionally
effects of confinement and external fields are investigated. The use of the
lattice Boltzmann approach allows to follow dynamics after the thermal
quench and electric field application. The
study is an important step in modeling of BPs and should deserve a
publication in Nature Communications. Nevertheless there are some weak
points that need improvement before the publication.
Below is the list of remarks."

RESPONSE:
We thank the reviewer for her/his generally encouraging comments, and for
recommending publication in Nature Communication subject to the resolution
of some points. Below are our detailed reply to each of the reviewer's
question, together with a description of how the comment has been dealt with
in the revision.

COMMENT:
"The term colloid is by the definition a dispersion of particles in a fluid.
Using the term also for the particles my lead to a confusion. In the
manuscript often also a more appropriate word
colloidal particle is used."

RESPONSE: We have standardised on "colloidal particle(s)" or simply
"particle(s)" where appropriate.

COMMENT:
"It would be important to add comments that would help experimentalist in
designing experiments with colloidal blue phases and perhaps stimulate
development of applications on the base of
these materials."

RESPONSE: The issue of experimental guidance is addressed in detail
in our reply to reviewer 2 below. In particular we have added a full
paragraph in the discussion (page 3, highlighted in red), with explicit 
reference to experimental studies on nanoparticle liquid crystal composites 
(currently Refs. [34-39]).

COMMENT:
"Page 2: In the discussion of the physical understanding of trapping of colloidal particles in BP disclination networks it would be useful to refer to the
initial study by Ravnik et al. in Faraday
Discussions 2009 devoted to this topic."

RESPONSE: Citation for Ravnik et al Faraday Discuss. 2010 (sic) is added
[current reference 16].

COMMENT: "250 K is more than the total mesophase stability range!"

RESPONSE: A stability range for templated blue phases of -125 to
+125 degrees Celsius is reported by Castles et al. Nat. Mat. 2012.
We have made this connection more clear by rewriting the relevant
sentence, which now reads (see page 2, highlighted in red): 
"However, recent advances in formulation have widened
this stability range enormously [11,12,13,14]. Most recently,
templated blue phases with a stability range of -125 to 125$^\circ$C
have been reported [Castles], paving the way to applications of BPs
in operational display devices."

COMMENT: The statement "This is defined such that W > 0 gives preferential
perpendicular orientation of the director field at the particle surface." is
not consistent with the interaction form
described in the Supplement.

RESPONSE: We agree this statement was confusing. It has been replaced by
a clearer statement (on page 4, highlighted in red): 
"The anchoring strength is defined such that
W > 0 imparts a preferred orientation to the director field at the
particle surface. This orientation can be either normal to the
surface or in a plane tangential to the surface, a property related
to surface chemistry in experimental systems but which can be imposed
at will in simulations."

COMMENT:
"For potential readers it would be useful to specify the particle radius in
nanometers in the main text. How it compares to the nematic correlation
length? How the radius influences the structures? How tiny the particles can 
still be that the method based on the Landau-de Gennes free energy is 
justified."

RESPONSE: We have added in the main text (last paragraph introduction) a
summary of the relevant dimensions in physical units:  "The simulations
presented here are representative of a liquid crystal blue phase with unit
cell size in the range of 100--500~nm and particles with a radius of
around 50~nm." This new paragraph also serves the requirement for a brief
summary of the results as set out in the manuscript checklist.

RESPONSE (cont.): As a coarse-grained description of the mesophase, the
Landau-de Gennes  theory is appropriate as long as the size of the
colloidal particle is much larger than the size of the mesogen,
the latter being typically in the range of 1--10 nanometres. Our
simulations meet this criteria (particle size about 50 nm) and the
use of Landau-de Gennes is therefore justified.

COMMENT: "Is WR/K = 2.8 justified for nanoparticles used in the study?"

RESPONSE: The choice of WR/K = 2.3 in the simulations (not 2.8)
is used to separate clearly a strong anchoring regime from a
weak anchoring regime (w = 0.23). While typical experimental
values for particles of this size may mean the weak regime is
more common, we do not see that this precludes interest in a
strong anchoring regime.

COMMENT: "Page 4:
Can the G elastic modulus be estimated for modeled structures?"

RESPONSE: We note we did have, in the original manuscript, a
paragraph which discusses estimates of the elastic modulus G
based on analogy with experimentally observed systems. To the
existing paragraph we have appended:
"This figure is consistent with an estimate obtained by comparing
the mechanical properties of the network of defects stabilised by
colloidal particles with those of rubber [Ramos et al 2002] for
the relevant simulation parameters (see Supplementary Information)."
To the supplementary information we append: "For a Frank elastic
constant of $K = 70$~pN and unit cell sizes in the range
$\lambda =$100--500~nm, a corresponding shear modulus $G \simeq
K/\lambda^2 =$0.3--7~kPa." The relevant reference is added.

COMMENT: "Page 5:
Can annealing of the structures obtained with the thermal quench lead to
ordered colloidal structures?"

RESPONSE: Our results suggest that this is probably not the case,
although it is difficult to systematically explore all the possible
quench protocol. That said, our data suggest that even starting with
colloidal particles randomly dispersed in an *ordered* blue phase
particles are able to distort the structure (see Fig. 1A, 1B); this
suggests that time evolution if anything may lead to more disorder,
rather than order.

COMMENT: "Is also in case of the disordered structures observed a change in the
stability range of the colloidal BPs?"

RESPONSE: Determining the stability range of the composite
phases was beyond the scope of this paper. This would have required 
an optimisation of the pitch length, which requires a large number 
of additional simulations for each individual parameter set.

COMMENT:
"In the discussion of colloidal particles in BP thin layers it would be good
to comment the study of Ravnik et al. in Soft Matter 2011 devoted 
the ordered structures of such a system."

RESPONSE: A citation for Ravnik et al (2011) has been added [current
reference 26].

COMMENT:
"Page 6: It would be good to comment the difference between structures in the 
bulk and thin layer for particles with strong anchoring. Does this depend on the
visualization or is more behind?"

RESPONSE: We have added the following at the end of the section
describing the confined geometry: "The degree of similarity between
the bulk and confined structures is somewhat masked by the local
change in particle density in the confined case. If one allows for
this, Figs.~1 and~2 show the individual structural motifs are actually
rather similar: templated lines for weak anchoring and clusters for
strong anchoring."

RESPONSE (cont.): One potential visualisation ambiguity is that the
displayed box sizes are different in each of the figures (cf. particle
size). An explicit statement has been added to each caption to make
this more clear to the reader. ADD CAPTIONS?

COMMENT:
"Are the electric field induced changes irreversible only in case of
colloidal dispersions?"

RESPONSE: For the given parameters the field induced changes are only
irreversible when colloids with sufficiently strong 
anchoring conditions are present. The mesophase without colloids 
undergoes a reversible ordering transition as described in reference [10].

COMMENT: "Page 7 Reference 2 has a mistake."

RESPONSE: The correct punctuation in the page numbers has been
inserted in the reference for Wright and Mermin [current reference 9].

COMMENT: "Supplement Segment 4.2:
The statement "Wall anchoring is either normal or planar with W=K = 1 in
simulation units" is unclear."

RESPONSE: This text has been replaced by a statement which is
transparent: "For both normal and planar anchoring at the walls,
the value of $W$ is the same as that used for the
strongly anchoring particles."

COMMENT:"References of the Supplement:
There are mistakes in references 1 and 2."

RESPONSE: Reference 1, Wright and Mermin, was correct in
the original supplement. Reference 2 has been updated to reflect
the 2nd edition of de Gennes and Prost, which now appears in the
main text [current reference 34].
CHECK REFERENCE

------------------------------------------------------------------------

Reply to Reviewer 2

COMMENT:
"This paper, "Self-assembly of colloid-cholesteric composites: A route to
switchable optical materials" reports simulations of the formation of
particle networks in a cholesteric host and in
particular in the blue phase of the cholesteric host. It reports the
influence of various physical parameters and in particular the surface
anchoring and concentration of the colloidal
particles. The paper also reports how external fields can be employed to
change the morphology of these colloidal composites and how this may be
employed in practical optical devices.

The paper builds on a body of previous research, both theoretical and
experimental that has explored the effects of dispersing particles in a
liquid crystal host. The novelty of the paper is
to model the effect of particles in the blue phases of a cholesteric host
ahd to postulate how these colloids may be employed in specific
applications.

I am not a theorteical physiscist and therefore will not comment on the
methodology but rather will take the perspective of an experimentalist and
how useful this paper is in understanding
these materials and more importantly how useful the paper is in suggesting
and designing experimental procedures and developing the actual materials.

The paper provides insight into the effectiveness of including particles
into the blue phase as a means of controlling and adjusting the structure
and stability of the materials. The paper
is clearly written and the figures, paricularly the included 3D videos,
provide clear perspective on the modeled phases of these materials. The
results are of interest and useful to experimentalists working in the field. "

RESPONSE:
We are grateful to the reviewer for her/his overall positive view of our
work. We also found her/his comments constructive and useful to improve
our paper. We have addressed them as follows.

COMMENT:
"The paper would be even more useful if it included references to actual 
materials, both the liquid crystal host and the added particles and perhaps
even suggested a model system that would be a good starting point to study
experimentally. "

RESPONSE: 
"We have fully taken this point on board.
In a new paragraph in the discussion section, we include six
new references specifically aimed at recent studies on experimental
nanoparticle liquid crystal composites. These are currently [34-39].
The paragraph also discusses the issue of a model system, and suggests
properties such a system may have to help match experiment and simulation."

COMMENT:
"In a similar vein, it would also be useful to address the
practical issue of reproducibility, which has been a major problem 
with these materials. The theoretical results clearly indicate how important 
the formation conditions are to the final result.

What is the authors sense of how easy it will be to reproducibly fabricate a 
device with a specific morphology and optical properties. This seems a likely
practical issue particularly with the rich variety of structures reported in
the simulations."

RESPONSE: Clearly the issue of experimental reproducibility is a difficult
one to be addressed by, of necessity idealised, simulation. However, one
point our simulations do show is that careful characterisation of not
only the particles and host fluid is required, but also of possible
confinement effects and associated anchoring conditions. This we also
address in our updated discussion, see the full paragraph highlighted in
red on pages 7-8.

Authors' note:

* ADDITIONAL CHANGES TO MEET FORMAT REQUIREMENTS
  cf Manuscript checklist

A number of other rearrangements have taken place to meet the
format requirements as set out the manuscript checklist.

Title

- The title has been adjusted to remove punctuation

Authors

- The full postal address has been added
- The corresponding author has been identified (Marenduzzo).

Abstract

- The abstract has been shorted to meet the 150 word limit.
  Citations have been removed to the main text.

Main text

- Symbols and units have been adjusted in places to come into style.

Results

- Subheadings of < 60 characters have been added

- References to supplementary items have been adjusted to come into style.

Methods

- All "methods" material (originally in supplementary information) is now
contained within the main paper. This includes relevant references.

- Dividing subheadings of < 60 characters have been added.

References

- Formatting is adjusted to come into style.

End notes

- A statement of author contributions has been added.

- A statement on conflicts of interest has been added.

Legends

- Figure legends have been adjusted to include a brief title.

Tables

- A title has been included.