### 2014-12-06 branches/kevin/docs/targetdp.tex
added targetdp notes
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2552 40876441-3496-0410-a613-a2fca35b2ae2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  targetdp.tex
%
%  Target Data Parallel information
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2014 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Target Data Parallel}

\subsection{Aim}

\subsection{Target context and API}

\subsection{Memory}

There are three types of memory requirement we will consider.
(A fourth, constant host memory not suitable for target constant
memory is could also be handled, but is less likely to occur.)
An approach to each relevant type is set out in the three following
sections.

\subsubsection{\texttt{const} object on host, and suitable for
\texttt{\_\_target\_const\_\_} memory}

Objects declared \texttt{const} on the host are suitable for transfer
to target constant memory. In order to have a consistent reference
to the object, the following approach can be used. We declare two
objects which are initialised via a macro-expansion at file scope:

\begin{lstlisting}
struct const_object {
  ...
};

#define object_initialiser() { ... }

const              struct const_object host = object_initialiser();
__target_const__ struct const_object target = object_initialiser();
\end{lstlisting}
All references to the object in kernel-related code is then via the
variable \texttt{target}, to be declared \texttt{extern} the the
relevant object header if necessary.
By making the appropriate macro-expansions for \texttt{\_\_target\_const\_\_}
we allow for different implementations. An implementation with separate
address spaces must ensure the \texttt{host}
contents are copied to the \texttt{target} object before any kernel
related activity. We suggest that this is at a well-defined
\begin{lstlisting}
__host__ int object_commit(struct object * obj);
\end{lstlisting}
stage. There is no coherency issue as both objects are constant. In a
shared-memory implementation, constant data are shared by virtue of
being at file scope.

We note a slightly more elegant version of the above would be
\begin{lstlisting}
struct const_object {
  /* declare */
};

const struct const_object object = {
  /* initialise */
};

const              struct const_object * const host = &object;
__target_const__ struct const_object * const target = &object;
\end{lstlisting}
which does away with the need for repeated initialisation via
an error-prone macro expansion. However,
for CUDA implementation the expansion of \texttt{\_\_target\_const\_\_}
to \texttt{\_\_constant\_\_}
may not get past \texttt{nvcc} unless an appropriate cast can
be identified to make the \texttt{target} initialisation consistent.

\subsubsection{Non \texttt{const} object on host, but suitable for
\texttt{\_\_target\_const\_\_} memory}

Here, the situation is less clear. Consider a non-constant data item
on the host defined as part of a object, along with corresponding
\texttt{target}:
\begin{lstlisting}
struct object {
   struct kernel_const host;
   ...
};

__target_const__ struct kernel_const target;

\end{lstlisting}
Here, an implementation with either uniform or separate address
spaces must arrange a copy of the data at appropriate times so
that any updates to \texttt{host} are correctly reflected
in \texttt{target}. All kernel-related references are again through
\texttt{target}.
The slightly disjoint nature of this definition and increased pressure
on global name-space must be weighed against potential increase in kernel
performance and reduction in register pressure.

\subsubsection{Non \texttt{const} object in both address spaces}

The most general case may be handled in the following manner:
\begin{lstlisting}

struct object_kernel_data {
  ...
};

struct object {
   struct object_kernel_data * host;
   struct object_kernel_data * target;
   ...
};

\end{lstlisting}
All relevant kernel data may then be handled by references to
\texttt{target} in a consistent way. For implementations with
separate address spaces, the memory coherency issue
must be handled explicitly as both \texttt{host} and \texttt{target}
are subject to modification. However, an implementation with uniform
address space may simply alias the \texttt{target} to the \texttt{host}
pointer. No data reside at file scope.

\subsubsection{Aliasing}
Consider a kernel data structure:
\begin{lstlisting}
struct object_kernel_data {
  double * data1;
  double * data2;
};
\end{lstlisting}
An object of this type is to be passed into a kernel function via a pointer,
which itself may be declared in contract to be
\begin{lstlisting}
__target__ void kernel_function(struct object_kernel_data * __restrict__ obj);
\end{lstlisting}
Further instruction to the compiler in the kernel function itself would
be required to specify the status of individual data items, e.g.,
\begin{lstlisting}
__target__ void kernel_function(struct object_kernel_data * __restrict__ obj) {

  double * restrict data1 = obj->data1;
  double * restrict data2 = obj->data2;

  /* ... and reference via ``data1'' and ``data2'' */
\end{lstlisting}
Standard C99 is required for a host implementation.
A test of this for a CUDA implementation with \texttt{nvcc} is needed.

\subsection{Kernel Execution}

\subsubsection{Typical kernel in host code}

A typical kernel representing a loop over local domain lattice sites
might look like:
\begin{lstlisting}

int nextra = 1;  /* extent of kernel into halo regions */
int nlocal[3];    /* local domain size */

for (ic = 1 - nextra; ic <= nlocal[X] + nextra; ic++) {
  for (jc = 1 - nextra; jc <= nlocal[Y] + nextra; jc++) {
    for (kc = 1 - nextra; kc <= nlocal[Z] + nextra; kc++) {

       /* operations for site (ic, jc, kc) which may involve neighbours */
       index = coords_index(ic, jc, kc);
       ...
    }
  }
}

\end{lstlisting}
Indexing calculations for lattice data structures are required to
respect the total extent of the halo regions. Not all sites are
involved in a typical kernel as determined above by \texttt{nextra}.

\subsubsection{Typical kernel in target code}

In a target implementation, the schematic code block shown above is necessarily
split into two parts. The first is the scheduling of the required
iterations (taking the place of the nested loops), and the second
is the site-wise kernel (the code inside the loops).

The two parts might take the form, schematically:
\begin{lstlisting}
int nextra = 1;            /* extent of kernel into halo regions */
int nlocal[3];             /* local domain size */
kernel_block_t nblocks;   /* Kernel blocks */
kernel_block_t ntpb;       /* Number of threads per block */

target_launch(kernel_function, nblocks, ntpb, nextra, nlocal[3], ...);
\end{lstlisting}
to be followed by
\begin{lstlisting}
__target__ void kernel_function(int nextra, int nlocal[3], ...) {

  if (threadIndex < threadsRequired) {

    /* locate (ic, jc, kc) from  kernel thread index */
    ...
  }

  return;
}

\end{lstlisting}
A small number of utility functions would allow us to compute the actual
number of threads required dependent on \texttt{nextra}, and to perform
relevant indexing. See, e.g., functions in
\texttt{branches/CUDA/src/utilities\_gpu.cu}
for identifying colloid map on the target.

Note that a \texttt{target\_launch()} macro expansion (that looks
like a standard C syntax) would require something of the form
\begin{lstlisting}
#define target_launch(f, nblocks, ntpb, ...) f(__VA_ARGS__)
\end{lstlisting}
or, for a CUDA run-time API implementation
\begin{lstlisting}
#define target_launch(f, nblocks, ntpb, ...) f<<<nblocks, ntpb>>>(__VA_ARGS__)
\end{lstlisting}

\subsubsection{Kernel reductions}

Can we do reductions in a kernel in a sane fashion?

\subsubsection{Bespoke target implementations for a single kernel}

What does the code look like if we are forced to have separate kernel
implementations for different targets for some reason?

\vfill
\pagebreak