### 2015-03-12 trunk/targetDP/doc/examples.tex

\chapter{Examples}\label{chapter:examples}
\begin{comment}
TO DO: describe this example. 
\end{comment}
\begin{figure}[h]
\begin{lstlisting}
__targetEntry__ void scale(double* t_field) { (*@\label{example:scalekernel:targetEntry}@*)

  int baseIndex;
  __targetTLP__(baseIndex, N) { (*@\label{example:scalekernel:targetTLP}@*)

    int iDim, vecIndex;
    for (iDim = 0; iDim < 3; iDim++) {

      __targetILP__(vecIndex)  \ (*@\label{example:scalekernel:targetILP}@*) 
         t_field[iDim*N + baseIndex + vecIndex] =  \
	 t_a*t_field[iDim*N + baseIndex + vecIndex];       	  
    }
  }
  return;
}
\end{lstlisting}
\caption{***}\label{fig:scalekernel}
\end{figure}

\begin{figure}[h]
\begin{lstlisting}
  targetMalloc((void **) &t_field, datasize); (*@\label{example:scalelaunch:targetMalloc}@*)

  copyToTarget(t_field, field, datasize); (*@\label{example:scalelaunch:copyToTarget}@*) 
  copyConstToTarget(&t_a, &a, sizeof(double)); (*@\label{example:scalelaunch:copyConstToTarget}@*)

  scale __targetLaunch__(N) (t_field); (*@\label{example:scalelaunch:targetLaunch}@*) 
  targetSynchronize();(*@\label{example:scalelaunch:targetSynchronize}@*)

  copyFromTarget(field, t_field, datasize); (*@\label{example:scalelaunch:copyFromTarget}@*) 
  targetFree(t_field); (*@\label{example:scalelaunch:targetFree}@*) 
\end{lstlisting}
\caption{***}\label{fig:scalelaunch}
\end{figure}