### 2015-03-12 trunk/targetDP/doc/introduction.tex

\chapter{Introduction}

\begin{comment}
TO DO.
\end{comment}
\section{Abstract}

**

\section{Motivation}

**

\section{Glossary}

\begin{itemize}

    \item \keyword{Accelerator}: ***
    \item \keyword{Host}: ***
    \item \keyword{Target}: ***
    \item \keyword{CUDA}: ***

\end{itemize}