### 2015-04-14 trunk/targetDP/doc/introduction.tex
more complete draft
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2632 40876441-3496-0410-a613-a2fca35b2ae2

\chapter{Introduction}

%\begin{comment}
%TO DO.
%\end{comment}
\section{Abstract}

The targetDP API provides an abstraction layer which allows
applications to \underline{t}arget \underline{D}ata
\underline{P}arallel hardware in a platform agnostic manner, by
abstracting the memory spaces and hierarchy of hardware
parallelism. Applications written using targetDP syntax are
performance portable: the same source code can be compiled for
different targets (currently supported are accelerators such as GPUs
and modern multi-core SIMD CPUs), without performance overheads. The
targetDP API is aimed at the types of parallelism found in grid-based
applications, but may be applicable to a wider class of problems. This
document introduces the targetDP memory and execution models, and
specifies the syntax and functionality of the
interface. Implementation details are also provided, and a simple
example is given to demonstrate usage.

\section{Motivation}
It is becoming increasingly difficult for applications to exploit
modern computers, which continue to increase in complexity and
diversity with features including multiple multicore CPUs (often
distributed across multiple nodes), vector floating point units,
accelerators such as GPUs and non-uniform distributed memory spaces.
From a scientist's perspective, it is not only imperative to achieve
performance, but also to retain maintainability, sustainability and
portability. The use of a simplistic, well structured and clearly
defined abstraction layer such as targetDP can allow the programmer to
express the scientific problems in a way that will automatically
achieve good performance across the range of leading hardware
solutions.

\section{Glossary}

\begin{itemize}

    \item \keyword{CPU}: Central Processing Unit. The main computer chip used in a system, suitable for a wide variety of computational tasks.
    \item \keyword{Accelerator}: A processing unit which is not used in isolation, but instead in tandem with the CPU, with the aim of improving the performance of key code sections. 
    \item \keyword{GPU}: Graphics Processing Unit. A type of accelerator, originally evolved to render graphics content (particularly to satisfy demands of the gaming industry), but now widely used for general purpose computation.
    \item \keyword{Host}: Another term for the CPU that ``hosts'' the application.
    \item \keyword{Data Parallel}: The type of algorithmic parallelism involved where a single operation is performed to each element of a data set. The extent of parallelism is determined by the size of the data set.
    \item \keyword{Target}: The device targeted for execution of data parallel operations.  Depending on the underlying hardware available, the target could simply be the a CPU, or it could be a separate device such as an accelerator.
    \item \keyword{CUDA}: Compute Unified Device Architecture. The parallel platform and model created by NVIDIA to allow general purpose programming of their GP architectures.
    \item \keyword{TLP}: Thread Level Parallelism.
    \item \keyword{ILP}: Instruction Level Parallelism.

\end{itemize}