### 2015-04-14 targetDP/doc/data_parallel_execution.tex
more complete draft
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@2632 40876441-3496-0410-a613-a2fca35b2ae2

\chapter{Data Parallel Execution}

\newpage
\section{targetEntry}
\subsection{Description}

The \verb+__targetEntry__+ keyword is used in a function declaration
or definition to specify that the function should be compiled for the
target, and that it will be called directly from host code.

\subsection{Syntax}
\begin{verbatim}
__targetEntry__ functionReturnType functionName(...
\end{verbatim}

\begin{itemize}
\item \verb+functionName+: The name of the function to be compiled for the target.
\item \verb+functionReturnType+: The return type of the function.
\item \verb+...+ the remainder of the function declaration or definition.
\end{itemize}

\subsection{Example}
See Line \ref{example:scalekernel:targetEntry} in Figure \ref{fig:scalekernel} in Section \ref{chapter:examples}.

\subsection{Implementation}
\subsubsection{C}
Holds no value.
\subsubsection{CUDA}
\verb+__global__+

\newpage
\section{target}
\subsection{Description}

The \verb+__target__+ keyword is used in a function declaration
or definition to specify that the function should be compiled for the
target, and that it will be called from a \verb+targetEntry+ or another \verb+target+ function.

\subsection{Syntax}
\begin{verbatim}
__target__ functionReturnType functionName(...
\end{verbatim}

\begin{itemize}
\item \verb+functionName+: The name of the function to be compiled for the target.
\item \verb+functionReturnType+: The return type of the function.
\item \verb+...+ the remainder of the function declaration or definition.
\end{itemize}

\subsection{Example}
Analogous to Line \ref{example:scalekernel:targetEntry} in Figure \ref{fig:scalekernel} in Section \ref{chapter:examples}.

\subsection{Implementation}
\subsubsection{C}
Holds no value.
\subsubsection{CUDA}
\verb+__device__+

\newpage
\section{targetHost}
\subsection{Description}

The \verb+__targetHost__+ keyword is used in a function declaration
or definition to specify that the function should be compiled for the
host.

\subsection{Syntax}
\begin{verbatim}
__targetHost__ functionReturnType functionName(...
\end{verbatim}

\begin{itemize}
\item \verb+functionName+: The name of the function to be compiled for the host.
\item \verb+functionReturnType+: The return type of the function.
\item \verb+...+ the remainder of the function declaration or definition.
\end{itemize}

\subsection{Example}
Analogous to Line \ref{example:scalekernel:targetEntry} in Figure \ref{fig:scalekernel} in Section \ref{chapter:examples}.

\subsection{Implementation}
\subsubsection{C}
Holds no value.
\subsubsection{CUDA}
\verb+extern ``C'' __host__+

\newpage
\section{targetLaunch}

\subsection{Description}

The \verb+__targetLaunch__+ syntax is used to launch a function across a data parallel target architecture.

\subsection{Syntax}
\begin{verbatim}
functionName __targetLaunch__(size_t extent) \
                 (functionArgument1,functionArgument2,...);
\end{verbatim}

\begin{itemize}
\item \verb+functionName+: The name of the function to be launched. This function must be declared as \verb+__targetEntry__+ .
\item \verb+functionArguments+: The arguments to the function \verb+functionName+  
\item \verb+extent+: The total extent of data parallelism.
\end{itemize}

\subsection{Example}
See Line \ref{example:scalelaunch:targetLaunch} in Figure \ref{fig:scalelaunch} in Section \ref{chapter:examples}.

\subsection{Implementation}
\subsubsection{C}
Holds no value.
\subsubsection{CUDA}
CUDA \verb+<<<...>>>+ syntax.

\newpage
\section{targetSynchronize}

\subsection{Description}

The \verb+targetSynchronize+ function is used to block until the preceding \verb+__targetLaunch__+ has completed.

\subsection{Syntax}
\begin{verbatim}
void targetSynchronize();
\end{verbatim}

\subsection{Example}
See Line \ref{example:scalelaunch:targetSynchronize} in Figure \ref{fig:scalelaunch} in Section \ref{chapter:examples}.
\subsection{Implementation}
\subsubsection{C}
Dummy function.
\subsubsection{CUDA}
\verb+cudaThreadSynchronize+

%% \newpage
%% \section{targetErrorCheck}

%% \subsection{Description}

%% The \verb+__targetErrorCheck__+ function is used to query for errors resulting from operations involving the target. The

%% \subsection{Syntax}
%% \begin{verbatim}
%% __targetSynchronize__()
%% \end{verbatim}

%% \subsection{Example}

%% \subsection{Implementation}
%% \subsubsection{C}

%% \subsubsection{CUDA}

\newpage
\section{targetTLP}\label{sec:TLP}

\subsection{Description}

The \verb+__targetTLP__+ syntax is used, within a \verb+__targetEntry__+
function, to specify that the proceeding block of code should be
executed in parallel and mapped to thread level parallelism
(TLP). Note that he behaviour of this operation depends on the defined
virtual vector length (VVL), which controls the lower-level
Instruction Level Parallelism (ILP) (see following section).

\subsection{Syntax}
\begin{verbatim}
__targetTLP__(int baseIndex, size_t extent) 
{
//code to be executed in parallel
}
\end{verbatim}

\begin{itemize}
\item \verb+extent+: The total extent of data parallelism, including
  both TLP and ILP
\item \verb+baseIndex+: the TLP index. This will vary from 0 to \verb+extent-VVL+ with stride VVL. This index should be combined with the ILP index to access shared arrays within the code block (see following section).

\end{itemize}

\subsection{Example}
See Line \ref{example:scalekernel:targetTLP} in Figure \ref{fig:scalekernel} in Section \ref{chapter:examples}.

\subsection{Implementation}
\subsubsection{C}
OpenMP parallel loop.
\subsubsection{CUDA}
CUDA thread lookup.

\newpage
\section{targetILP}\label{sec:ILP}

\subsection{Description}

The \verb+__targetILP__+ syntax is used, within a \verb+__targetTLP__+
region, to specify that the proceeding block of code should be
executed in parallel and mapped to instruction level parallelism
(ILP), where the extent of the ILP is defined by the virtual vector length (VVL) in the targetDP implementation. 
\begin{comment}
need to document VVL in more detail somewhere and link.
\end{comment}

\subsection{Syntax}
\begin{verbatim}
__targetILP__(int vecIndex) 
{
//code to be executed in parallel
}
\end{verbatim}

\begin{itemize}
\item \verb+baseIndex+: the ILP index. This will vary from 0 to \verb+VVL-1+.  This index should be combined with the TLP index to access shared arrays within the code block (see previous section).
\end{itemize}

\subsection{Example}
See Line \ref{example:scalekernel:targetILP} in Figure \ref{fig:scalekernel} in Section \ref{chapter:examples}.

\subsection{Implementation}
\subsubsection{C}
Short vectorizable loop.
\subsubsection{CUDA}
Short vectorizable loop.

\newpage
\section{targetCoords3D}

\subsection{Description}

The \verb+targetCoords3D+ function provides the 3D lattice coordinates corresponding to a specified linear index.

\subsection{Syntax}
\begin{verbatim}
void targetCoords3D(int coords3D[3], int extent3D[3], int index);
\end{verbatim}

\begin{itemize}
\item \verb+coords3D+ (output): an array of 3 integers to be populated with the 3D coordinates.
\item \verb+extent3D+ (input): An array of 3 integers corresponding to the 3D dimensions of the lattice.
\item \verb+index+ (input): the linear index.

\end{itemize}

\subsection{Example}
**TO DO**
\subsection{Implementation}
\subsubsection{C}
**TO DO**
\subsubsection{CUDA}
**TO DO**

\newpage
\section{targetIndex3D}

\subsection{Description}

The \verb+targetIndex3D+ function returns the linear index corresponding to a specified set of 3D lattice coordinates.

\subsection{Syntax}
\begin{verbatim}
int targetIndex3D(int Xcoord,int Ycoord,int Zcoord,int extent3D[3]);
\end{verbatim}

\begin{itemize}
\item \verb+Xcoord+ (input): the specified coordinate in the X direction.
\item \verb+Ycoord+ (input): the specified coordinate in the Y direction.
\item \verb+Zcoord+ (input): the specified coordinate in the Z direction.
\item \verb+extent3D+ (input): an array of 3 integers corresponding to the 3D dimensions of the lattice.
\end{itemize}

\subsection{Example}
**TO DO**
\subsection{Implementation}
\subsubsection{C}
**TO DO**
\subsubsection{CUDA}
**TO DO**