### 2015-05-14 incite/Milestone_Table_template.tex
adding incite bid templates
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@2645 40876441-3496-0410-a613-a2fca35b2ae2

%\documentclass[12pt,letterpaper,english]{article}
\documentclass[11pt,landscape]{article}
\usepackage[T1]{fontenc} % Standard package for selecting font encodingsamely titles, headers and contents
\usepackage{txfonts} % makes spacing between characters space correctly
\usepackage{xcolor} % Driver-independent color extensions for LaTeX and pdfLaTeX.
%\usepackage[pagestyles,raggedright]{titlesec} % related with sections—n
%\usepackage{blindtext} % To create text
\usepackage{fancyhdr} % header footer placement

\usepackage[top=1in, bottom=1in, left=1in, right=1in] {geometry} % Margins
\usepackage{graphicx}  % Essential for adding images to you document.

\usepackage{sectsty}
\sectionfont{\large}
\subsectionfont{\normalsize}
\subsubsectionfont{\normalsize \it}

\usepackage{caption}
\captionsetup{labelsep=period}

\renewcommand{\arraystretch}{1.25}

\pagestyle{fancy} % allows you to use the header and footer commands

\raggedright
\begin{document}

\setlength{\parindent}{0in} % Amount of indentation at the first line of a paragraph.

\pagestyle{fancy} \lhead{Title} \rhead{Lead PI} \renewcommand{%
\headrulewidth}{0.0pt}

\colorbox{yellow} {\bf {\emph{[Refer to the guidelines for instructions in preparing the proposal. Table does not count toward}}}
\colorbox{yellow} {\bf {\emph{[15-page project narrative count.]}}}

\bigskip

\textbf{Proposal Title (exactly as it appears on submission):} Text

\begin{table}[h]
%\centering
%\caption{Table title}
%\vspace{-.15in}
\label{Tab1}
\begin{tabular}{p{5cm} p{5cm} p{5cm}} \\ \hline 
\noalign{\smallskip}
\multicolumn{3}{l}{\textbf {Year 1}}\\ \hline 
\noalign{\smallskip}
\textbf{Milestone} & \textbf{Details} & \textbf{Date and Status} \\ \hline 
Insert Text & Insert Text & Insert Text \\ \hline 
Insert Text & Insert Text & Insert Text \\ \hline 
Insert Text & Insert Text & Insert Text \\ \hline 
\noalign{\smallskip}
\multicolumn{3}{l}{\textbf {Year 2 (if appropriate)}}\\ \hline
\noalign{\smallskip}
Insert Text & Insert Text & Insert Text \\ \hline 
Insert Text & Insert Text & Insert Text \\ \hline 
Insert Text & Insert Text & Insert Text \\ \hline 
\noalign{\smallskip}
\multicolumn{3}{l}{\textbf {Year 3 (if appropriate)}}\\ \hline
\noalign{\smallskip}
Insert Text & Insert Text & Insert Text \\ \hline 
Insert Text & Insert Text & Insert Text \\ \hline 
Insert Text & Insert Text & Insert Text \\ \hline

\end{tabular}
\end{table}

\end{document}