### 2015-05-14 incite/Project_Summary_template.tex

\documentclass[11pt,letterpaper,english]{article}
\usepackage[T1]{fontenc} % Standard package for selecting font encodingsamely titles, headers and contents
\usepackage{txfonts} % makes spacing between characters space correctly
\usepackage{xcolor} % Driver-independent color extensions for LaTeX and pdfLaTeX.
%\usepackage[pagestyles,raggedright]{titlesec} % related with sections—n
%\usepackage{blindtext} % To create text
\usepackage{fancyhdr} % header footer placement

\usepackage[top=1in, bottom=1in, left=1in, right=1in] {geometry} % Margins
\usepackage{graphicx}  % Essential for adding images to you document.

\usepackage{sectsty}
\sectionfont{\normalsize}
\subsectionfont{\normalsize}
\subsubsectionfont{\normalsize \it}

\usepackage{caption}
\captionsetup{labelsep=period}

\pagestyle{fancy} % allows you to use the header and footer commands

\raggedright
\begin{document}

\setlength{\parindent}{0in} % Amount of indentation at the first line of a paragraph.

\pagestyle{fancy} \lhead{Title} \rhead{Lead PI} \renewcommand{%
\headrulewidth}{0.0pt}

\begin{center}
\bf {PROJECT EXECUTIVE SUMMARY} \\
{\bf \small {\em (Must not exceed 1 page but does} not {\em count toward the 15-page project narrative limit.)}}
\end{center}

%\begin{flushleft}
\bigskip

\textbf{Title (\emph{80 characters max; strictly enforced})}: Text \smallskip

\textbf{PI and Co-PI(s)}: Text \smallskip

\textbf{Applying Institution/Organization}: Text \smallskip

\textbf{Number of Processor Hours Requested}: Text \smallskip

\textbf{Amount of Storage Requested}: Text \smallskip

\textbf{Executive Summary ({\emph{May use the remainder of page}}):} \\

The executive summary should accurately describe the proposed research and the high-impact scientific or technical advances you will realize with the proposed INCITE allocation. Industry organizations should also summarize the potential economic or strategic business impact of the proposed research.\\
\vspace{.15in}
Insert paragraph(s).

%\end{flushleft}

\end{document}