### 2015-05-14 incite/Publications_template.tex

\documentclass[11pt,letterpaper,english]{article}
\usepackage[T1]{fontenc} % Standard package for selecting font encodings
\usepackage{txfonts} % makes spacing between characters space correctly
\usepackage{xcolor} % Driver-independent color extensions for LaTeX and pdfLaTeX.
%\usepackage[pagestyles]{titlesec} % related with sections—namely titles, headers and contents
%\usepackage{blindtext} % To create text
%\usepackage{mdwlist} % mdwlist for compact enumeration/list items 
\usepackage{fancyhdr} % header footer placement

\usepackage[top=1in, bottom=1in, left=1in, right=1in] {geometry} % Margins
\usepackage{graphicx}  % Essential for adding images to you document.

\usepackage{sectsty}
\sectionfont{\normalsize}
\subsectionfont{\normalsize}
\subsubsectionfont{\normalsize \it}

\usepackage{caption}
\captionsetup{labelsep=period}

\setlength{\parskip}{\baselineskip}%
\setlength{\parindent}{0pt}%

\pagestyle{fancy} % allows you to use the header and footer commands

\raggedright
\begin{document}

\setlength{\parindent}{0in} % Amount of indentation at the first line of a paragraph.

\pagestyle{fancy} \lhead{Title} \rhead{Lead PI} \renewcommand{%
\headrulewidth}{0.0pt}

\begin{center}
\bf {PUBLICATIONS RESULTING FROM INCITE AWARDS} \\
{\bf {\em (Does} not {\em count toward the page limit.)}}
\end{center}
\vspace{-.25in}
Provide a list of publications resulting from previous INCITE awards to this project team for work related to the proposal under consideration. Only those publications that include an acknowledgement to INCITE and/or the LCF may be included. This list may {\bf \underline {not}} be used in lieu of references in the project narrative. If applicable, list the citation both here and in the project narrative.

References must be single-column format, 11 point, Arial or Times New Roman.

%{\noindent \bf {REFERENCE EXAMPLES}}

%\leftskip 0.25in
% \parindent -0.25in
%{Author, First, Second Author, and Third Author, ``An article in a journal,'' {\em{Journal Name}} {\bf{32}}(4): 46--52.}\\
%{Author, First, and Second Author, {\em{Report Title}}, Report Number, Publishing Organization or Agency, City, State, 2013.}\\
%{Author, First, et al., ``Chapter Title,'' {\em{Book Title}}, Publisher, City, State, 2011.}\\
%{Corporate or Agency Author, {\em{Book Title}}, Publisher, City, State, 2012.}

\end{document}