### 2015-05-14 trunk/incite/CV_template.tex
adding incite bid templates
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2645 40876441-3496-0410-a613-a2fca35b2ae2

\documentclass[11pt,letterpaper,english]{article}
\usepackage[T1]{fontenc} % Standard package for selecting font encodings
\usepackage{txfonts} % makes spacing between characters space correctly
\usepackage{xcolor} % Driver-independent color extensions for LaTeX and pdfLaTeX.
%\usepackage{blindtext} % To create text
%\usepackage{mdwlist} % mdwlist for compact enumeration/list items 
%\usepackage[pagestyles]{titlesec} % related with sections—namely titles, headers and contents
\usepackage{fancyhdr} % header footer placement

\usepackage[top=1in, bottom=1in, left=1in, right=1in] {geometry} % Margins
\usepackage{graphicx}   % Essential for adding images to you document.

\usepackage{sectsty}
\sectionfont{\large}
\subsectionfont{\normalsize}
\subsubsectionfont{\normalsize \it}

\usepackage{caption}
\captionsetup{labelsep=period}

\pagestyle{fancy} % allows you to use the header and footer commands

\raggedright
\begin{document}

\setlength{\parindent}{0in} % Amount of indentation at the first line of a paragraph.

\pagestyle{fancy} \lhead{Title} \rhead{Lead PI} \renewcommand{%
\headrulewidth}{0.0pt}

\centering {\bf Curriculum Vitae ({\emph{2-page limit}})}\\
{\bf PI NAME}\\
{\bf Contact Information} \smallskip

\begin{flushleft} {\bf Professional Preparation}
{\parindent 16pt

PhD \\ 
MS \\ 
BS \\ 
}

\vspace{.04in}
{\bf Appointments}
{\parindent 16pt

2010--present \\ 
2001--2010 \\ 
1990--2001 \\ 
}

\vspace{.04in}
{\bf Five Publications Most Relevant to This Proposal}
\vspace{-6pt}
\begin{enumerate} \itemsep1pt \parskip0pt \parsep0pt
\item Publication most relevant to this proposal \\ 
\item Publication most relevant to this proposal \\ 
\item Publication most relevant to this proposal \\ 
\item Publication most relevant to this proposal \\ 
\item Publication most relevant to this proposal \\ 
\end{enumerate}

\vspace{-6pt}
{\bf Research Interests and Expertise}
{\parindent 16pt

Insert paragraph(s).
}

\vspace{.04in}
{\bf Synergistic Activities}
\vspace{-6pt}
\begin{enumerate} \itemsep1pt \parskip0pt \parsep0pt
\item Text \\ 
\item Text \\ 
\item Text \\ 
\item Text \\ 
\item Text \\ 
\end{enumerate}

\vspace{-6pt}
{\bf Collaborators ({\emph{past 5 years including name and current institution}})} 
{\parindent 16pt

Collaborator name, current institution \\
Collaborator name, current institution \\
}

\end{flushleft}

\end{document}