### 2015-10-23 trunk/docs/develop.tex
Issue 1405: added new docs directory with content
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2764 40876441-3496-0410-a613-a2fca35b2ae2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  develop.tex
%
%  Further information for delvelopers
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2014-2015 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Further Information for Developers}

\subsection{Address models for 3-dimensional fields}

The code allows for different order of storage associated with 3-dimensional
fields (scalars, vectors, and so on). For historical reasons these are
referred to as `Model' and `Model~R' options, which correspond
to array-of-structures and structure-of-arrays layout, respectively. Early
versions of the code for CPU were written to favour summation of LB
distributions on a per lattice site basis in operations such as
$\rho(\mathbf{r}) = \sum_i f_i(\mathbf{r})$. This is array-of-structures,
where the $f_i$ are stored contiguously per site. Introduction of GPU
versions, where memory coalescing favours the opposite memory order,
were then referred to as `Model~R', the `R' standing for `reverse'.

The memory layouts are discussed below. In all cases, a 3-d field occupies
lattice sites with a unique spatial index determine by position, and
computed via \texttt{coords\_index()}. These position indices will be denoted
$r_0, r_1, \ldots, r_{n_s-1}$ where $n_s$ is the total number of lattice
sites (including halo points).

\subsubsection{Rank 1 objects (to include scalar fields)}

\textbf{ADDR\_MODEL}: The array-of-structures or Model order for an
$n-$vector field
with components $ v_\alpha = (v_0, v_1, \ldots, v_{n-1})$ is, schematically:
\[
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_0}
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_1}
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_2} \ldots
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_{n_s-1}}
\]

\textbf{ADDR\_MODEL\_R}: The Model~R version is:
\[
\overbrace{
\begin{array}{cc} \boxed{v_0} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_0} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_0}} \\ \phantom{r_2}\end{array} \mkern-22mu
\begin{array}{cc} \boxed{v_0} \\  _{r_{n_s-1}} \end{array}
}^{component_0}
\overbrace{
\begin{array}{cc} \boxed{v_1} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_1} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_0}} \\ \phantom{r_2}\end{array} \mkern-22mu
\begin{array}{cc} \boxed{v_1} \\  _{r_{n_s-1}} \end{array}
}^{component_1}
\begin{array}{cc} \ldots \\ \phantom{r_2} \end{array}
\overbrace{
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_{n-1}}} \\ \phantom{r_2}\end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_{n_s-1}} \end{array}
}^{component_{n-1}}
\]

A scalar field has $n=1$.

% ADDR_MACRO(n_s, n, i, \alpha) in this notation corresponds to
% ADDR_MACRO(nsites, nfield, index, ifld)

\subsubsection{Rank 2 objects (to include dyadic tensor fields)}

\textbf{ADDR\_MODEL}:
A general rank 2 tensor $t_{\alpha\beta}$ with components $(t_{0,0},
\ldots, t_{m-1,n-1})$ is stored as:
\begin{gather*}
\underbrace{
\boxed{t_{0,0}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,0}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,0}\vphantom{t_{m-1,n-1}}}
\boxed{t_{0,1}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,1}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,1}\vphantom{t_{m-1,n-1}}}
\ldots
\boxed{t_{0,n-1}\vphantom{t_{m-1,n-1}}}\boxed{t_{1,n-1} \vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,n-1}\vphantom{t_{m-1,n-1}}}
}_{r_0}
\ldots
\\
\ldots
\underbrace{
\boxed{t_{0,0}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,0}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,0}\vphantom{t_{m-1,n-1}}}
\boxed{t_{0,1}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,1}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,1}\vphantom{t_{m-1,n-1}}}
\ldots
\boxed{t_{0,n-1}\vphantom{t_{m-1,n-1}}}\boxed{t_{1,n-1} \vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,n-1}\vphantom{t_{m-1,n-1}}}
}_{r_{n_s-1}}
\end{gather*}
Dyadic tensors, for example the gradient of a vector field
$\partial_\alpha v_\beta$ in three dimensions, are stored in corresponding
fashion with $m=3$ and
$\partial_\alpha = (\partial_x, \partial_y, \partial_z)$.

\textbf{ADDR\_MODEL\_R}: The Model R version is
\[
\overbrace{
\begin{array}{cc} \boxed{t_{0,0}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{0,0}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{t_{0,0}}} \\ \phantom{r_2}\end{array} \mkern-18mu
\begin{array}{cc} \boxed{t_{0,0}} \\  _{r_{n_s-1}} \end{array}
}^{component_{0,0}}
\overbrace{
\begin{array}{cc} \boxed{t_{1,0}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{1,0}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{t_{0,0}}} \\ \phantom{r_2}\end{array} \mkern-18mu
\begin{array}{cc} \boxed{t_{1,0}} \\  _{r_{n_s-1}} \end{array}
}^{component_{1,0}}
\begin{array}{cc} \ldots\vphantom{t_{n-1}} \\ \phantom{r_2}\end{array}
\overbrace{
\begin{array}{cc} \boxed{t_{m-1,n-1}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{m-1,n-1}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{t_{0,0}}} \\ \phantom{r_2}\end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{m-1,n-1}} \\  _{r_{n_s-1}} \end{array}
}^{component_{m-1,n-1}}
\]

% ADDRRES_MACRO(n_s, m, n, i, alpha, beta) corresponds to, e.g.,
% ADDRESS_MACRO(nsites, NVECTOR, NQAB, index, ia, ib)

\subsubsection{Compressed rank 2 objects}

A symmetric tensor $S_{\alpha\beta}$ in three dimensions has six independent
components. It may be convenient to store this in compressed form as a
rank 1 vector $(S_{xx}, S_{xy}, S_{xz}, S_{yy}, S_{yz}, S_{zz})$ to eliminate
redundent storage.

A symmetric traceless rank 2 tensor --- for example, the Landau-de Gennes
liquid crystal order parameter $Q_{\alpha\beta}$ --- has five independent
components. This is stored as a rank 1 vector with five components
$(Q_{xx}, Q_{xy}, Q_{xz}, Q_{yy}, Q_{yz})$ to eliminate redundant
storage. API calls are provided to expand the compressed format to the
full rank-2 representation $Q_{\alpha\beta}$ and, conversely, to compress
the full representation to five components.

\subsubsection{Rank 3 objects (to include triadic tensor fields)}

The general rank 3 object $t_{\alpha\beta\gamma}$ with components
$(t_{0,0,0}, \ldots, t_{m-1,n-1,p-1})$ is stored in a manner which
generalises from the above, i.e., with the rightmost index running
fastest. Diagrams are omitted, but Model and Model R storage patterns
follow the same form as seen above.

A triadic tensor, for example the general second derivative of a vector
field $\partial_\alpha \partial_\beta v_\gamma$ may be stored as a rank
3 object.

\subsubsection{Compressed rank 3 objects}

Symmetry properties may be used to reduce the storage requirement
associated with rank 3 tensors. `For example, the gradient of the
liquid crystal order parameter $\partial_\gamma Q_{\alpha\beta}$
may be stored as a rank 2 object. The exact requirement will depend
on the construction of the tensor.

%
% The generalised macro would be
% ADDR_MACRO(nsites, m, n, p, i, alpha, beta, gamma)

\subsubsection{LB distribution data}

\textbf{ADDR\_MODEL}: For the LB distributions, up to two distinct
distributions can be accommodated to allow for a binary fluid
implementation, although the usual situation is to have only one.
The Model order for two $N$-velocity distributions $f$ and $g$ is:
\[
\underbrace{
\boxed{f_0\vphantom{f_{N-1}}}    \boxed{f_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{f_{N-1}} \,
\boxed{g_0\vphantom{f_{N-1}}}        \boxed{g_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{g_{N-1}\vphantom{f_{N-1}}}}_{r_0}
\ldots
\underbrace{
\boxed{f_0\vphantom{f_{N-1}}}    \boxed{f_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{f_{N-1}} \,
\boxed{g_0\vphantom{f_{N-1}}}        \boxed{g_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{g_{N-1}\vphantom{f_{N-1}}}}_{r_{n_s=1}}
\]

\textbf{ADDR\_MODEL\_R}: The Model R order is:

\[
\overbrace{
\overbrace{
\begin{array}{cc} \boxed{f_0} \\  _{r_0} \end{array} \mkern-18mu
\begin{array}{cc} \boxed{\ldots\vphantom{f_0}} \\ \phantom{r_2}\end{array} \mkern-22mu
\begin{array}{cc} \boxed{f_0} \\  _{r_{n_s-1}} \end{array}
}^{f_0}
\begin{array}{cc} \ldots \\ \phantom{r_2} \end{array}
\overbrace{
\begin{array}{cc} \boxed{f_{N-1}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{f_{N-1}}} \\ \phantom{r_2}\end{array} \mkern-17mu
\begin{array}{cc} \boxed{f_{N-1}} \\  _{r_{n_s-1}} \end{array}
}^{f_{N-1}}
}^{f_i}
\overbrace{
\overbrace{
\begin{array}{cc} \boxed{g_0\vphantom{f_{N-1}}} \\ _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{f_{N-1}}} \\ \phantom{r_2}\end{array} \mkern-22mu
\begin{array}{cc} \boxed{g_0\vphantom{f_{N-1}}} \\  _{r_{n_s-1}} \end{array}
}^{g_0} \ldots
\overbrace{
\begin{array}{cc} \boxed{g_{N-1}\vphantom{f_{N-1}}} \\ _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{f_{N-1}}} \\ \phantom{r_2}\end{array} \mkern-17mu
\begin{array}{cc} \boxed{g_{N-1}\vphantom{f_{N-1}}} \\  _{r_{n_s-1}} \end{array}
}^{g_{N-1}\vphantom{f_{N-1}}}
}^{g_i} 
\]
For the single-distribution case, this is equivalent to the rank 1 vector
field with $n=N$.

\vfill
\pagebreak