### 2016-01-12 branches/kevin-1397/targetDP/doc/examples.tex
Add eature branch
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2811 40876441-3496-0410-a613-a2fca35b2ae2

\chapter{Example}\label{chapter:examples}

Consider a simple example: the scaling of a 3-vector
field by a constant, as implemented in a sequential programming style in Figure \ref{fig:scaleorig}. On each lattice site exists a 3-vector (a collection of three values corresponding to the 3 spatial dimensions). The outer loop corresponds to lattice sites, and the inner loop to the 3 components within each lattice site. This is a simple example of operations on ``multi-valued'' data, a very common situation in scientific simulations.

The lattice-based parallelism corresponding to the outer loop can be
mapped to data parallel hardware using targetDP. We introduce targetDP
by replacing the sequential code with the function given in Figure
\ref{fig:scalekernel}. The \verb+t_+ syntax is used to identify target
data structures. The \verb+__targetEntry__+ syntax is used to specify
that this function is to be executed on the target, and it will be
called from host code.  We expose the lattice-based parallelism to
each of the TLP and ILP levels of hardware parallelism through use of
the combination \verb+__targetTLP__(baseIndex,N)+ and
\verb+__targetILP__(vecIndex)+ (See Sections \ref{sec:TLP} and
\ref{sec:ILP}). The former specifies that lattice-based parallelism
should be mapped to TLP, where each thread operates on a chunk of
lattice sites. The latter specifies that the sites within each chunk
should be mapped to ILP. It can be seen that the \verb+t_field+ array
is accessed by combining these indexes. The size of the chunk can be
set within the targetDP implementation, to give the best performance
for a particular architecture.

The \verb+scale+ function is called from host code as shown in Figure \ref{fig:scalelaunch}. The memory management facilities are used to allocate and transfer data to and from the target, as described in Chapter \ref{chap:memmanage}.

\begin{figure}[h]
\begin{lstlisting}
 for (idx = 0; idx < N; idx++) { //loop over lattice sites
   int iDim;
   for (iDim = 0; iDim < 3; iDim++) 
      field[iDim*N+idx] = a*field[iDim*N+idx];
 }

\end{lstlisting}
\caption{A sequential implementation of the scalar multiplication of each element of a lattice data structure.}\label{fig:scaleorig}
\end{figure}

\begin{figure}[h]
\begin{lstlisting}
__targetEntry__ void scale(double* t_field) { (*@\label{example:scalekernel:targetEntry}@*)

  int baseIndex;
  __targetTLP__(baseIndex, N) { (*@\label{example:scalekernel:targetTLP}@*)

    int iDim, vecIndex;
    for (iDim = 0; iDim < 3; iDim++) {

      __targetILP__(vecIndex)  \ (*@\label{example:scalekernel:targetILP}@*) 
         t_field[iDim*N + baseIndex + vecIndex] =  \
	 t_a*t_field[iDim*N + baseIndex + vecIndex];       	  
    }
  }
  return;
}
\end{lstlisting}
\caption{The targetDP implementation of the scalar multiplication kernel.}\label{fig:scalekernel}
\end{figure}

\begin{figure}[h]
\begin{lstlisting}
  targetMalloc((void **) &t_field, datasize); (*@\label{example:scalelaunch:targetMalloc}@*)

  copyToTarget(t_field, field, datasize); (*@\label{example:scalelaunch:copyToTarget}@*) 
  copyConstToTarget(&t_a, &a, sizeof(double)); (*@\label{example:scalelaunch:copyConstToTarget}@*)

  scale __targetLaunch__(N) (t_field); (*@\label{example:scalelaunch:targetLaunch}@*) 
  targetSynchronize();(*@\label{example:scalelaunch:targetSynchronize}@*)

  copyFromTarget(field, t_field, datasize); (*@\label{example:scalelaunch:copyFromTarget}@*)

  targetFree(t_field); (*@\label{example:scalelaunch:targetFree}@*) 
\end{lstlisting}
\caption{The host code used to invoke the targetDP scalar multiplication kernel.}\label{fig:scalelaunch}
\end{figure}