### 2016-06-06 branches/Anthony-nonblocking/targetDP/doc/introduction.tex
non-blocking halo exchange version
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2897 40876441-3496-0410-a613-a2fca35b2ae2

\chapter{Introduction}

%\begin{comment}
%TO DO.
%\end{comment}

It is becoming increasingly difficult for applications to exploit
modern computers, which continue to increase in complexity and
diversity with features including multicore/manycore CPUs, vector
floating point units, accelerators such as GPUs and non-uniform
distributed memory spaces.  From a scientist's perspective, it is not
only imperative to achieve performance, but also to retain
maintainability, sustainability and portability. The use of a
simplistic, well structured and clearly defined abstraction layer such
as targetDP can allow the programmer to express the scientific
problems in a way that will automatically achieve good performance
across the range of leading hardware solutions.

The targetDP API (first introduced in \cite{hpcc}) provides an
abstraction layer which allows applications to \underline{t}arget
\underline{D}ata \underline{P}arallel hardware in a platform agnostic
manner, by abstracting the memory spaces and hierarchy of hardware
parallelism. Applications written using targetDP syntax are
performance portable: the same source code can be compiled for
different targets (where we currently support GPU accelerators and
modern multicore or manycore SIMD CPUs), without performance
overheads. The model is appropriate for abstracting the parallelism
contained within each compute node, and can be combined with, e.g. MPI
to allow use on systems containing multiple nodes. The targetDP API is
primarily aimed at the types of parallelism found in grid-based
applications, but may be applicable to a wider class of problems.

The targetDP memory and execution models are described in Chapters \ref{sec:exmodel} and \ref{sec:memmodel} respectively, and the document goers on to specify the memory management functionality in Chapter \ref{chap:memmanage} and data parallel execution functionality in Chapter  \ref{sec:dpexecution}. Implementation details are also provided for the existing C and CUDA versions of targetDP throughout these chapters. Finally, a simple
example is given to demonstrate usage in Chapter \ref{chapter:examples}.

\section*{Glossary}

\begin{itemize}

    \item \keyword{CPU}: Central Processing Unit. The main computer chip used in a system, suitable for a wide variety of computational tasks.
    \item \keyword{Accelerator}: A processing unit which is not used in isolation, but instead in tandem with the CPU, with the aim of improving the performance of key code sections. 
    \item \keyword{GPU}: Graphics Processing Unit. A type of accelerator, originally evolved to render graphics content (particularly to satisfy demands of the gaming industry), but now widely used for general purpose computation.
    \item \keyword{Host}: Another term for the CPU that ``hosts'' the application.
    \item \keyword{Data Parallel}: The type of algorithmic parallelism involved where a single operation is performed to each element of a data set. The extent of parallelism is determined by the size of the data set.
    \item \keyword{Target}: The device targeted for execution of data parallel operations.  Depending on the underlying hardware available, the target could simply be the a CPU, or it could be a separate device such as an accelerator.
    \item \keyword{CUDA}: Compute Unified Device Architecture. The parallel platform and model created by NVIDIA to allow general purpose programming of their GPU architectures.
    \item \keyword{TLP}: Thread Level Parallelism.
    \item \keyword{ILP}: Instruction Level Parallelism.

\end{itemize}